---
title: "Visualizing and Summarizing Wildfire Damage in California"
author: "Sardys Avile-Martinez"
date: "2025-04-30"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing and Summarizing Wildfire Damage in California}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(wildvizR)
library(dplyr)
library(ggplot2)
library(knitr)
```

#  Introduction
Welcome to wildvizR — a lightweight R package that empowers users to explore, summarize, and visualize wildfire data in California. This vignette demonstrates how to use the core functions of the package with a sample dataset included in the package folder.

Let’s walk through it.

#  Load the Data
The dataset is stored in inst/extdata and accessed with system.file():

```{r read-data, echo=TRUE}

wildfire_data <- read.csv(system.file("extdata", "Cleaned_California_Wildfire_Damage.csv", package = "wildvizR"))
wildfire_data <- clean_column_names(wildfire_data)  
head(wildfire_data)

```

# Summarize Wildfire Damage by Year
Use summarize_wildfire_damage() to get an annual summary of acres burned and total financial loss.
```{r}

summary <- summarize_wildfire_damage(wildfire_data)
print(summary)

```

# Identify Major Incidents
This function flags wildfires with losses greater than $50 million by default:
```{r}
major <- flag_major_incidents(wildfire_data)
head(major)

```

# Inspect Data Structure and Missing Value
Use this utility to quickly review your dataset:
```{r}
inspect_data(wildfire_data)

```

# Plot County-Level Loss
Generate a choropleth map showing estimated financial loss by county:
```{r}
plot_county_loss_map(wildfire_data)

```







